---
title: "WQD7005_GroupProject_EDA"
author: "LOO JIA HAO (17218692)"
date: "5/16/2021"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) 
library(ggcorrplot)
library(ggplot2)
library(cowplot)

df <- read.csv("WA_Fn-UseC_-Telco-Customer-Churn.csv", stringsAsFactors = TRUE)

theme_bar <- theme_bw() +
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.position='none')


theme_pie <- theme_void() +
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        axis.text = element_text(size = 12),
        legend.position='none')

theme_donut <- theme_void() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position='none')
        
# '#F8766D', '#00BA38', '#619CFF'

theme_2 <- theme_bw()+
theme(axis.text.x = element_text(angle = 0, hjust = 1, vjust = 0.5), legend.position='none')

theme_3 <- theme_bw()+
theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position='none')

```

# 4. Exploratory Analysis
_TODO_
In the SAS SEMMA of Explore , this stage further discovery the data by identifying for hidden trends, important information and anomalies in order to acquire comprehension and insight

```{r}
df$SeniorCitizen <- as.factor(ifelse(df$SeniorCitizen==1, 'Yes', 'No'))
```

## 4.1. Target Distribution

First, let's check the distribution of our target variable, the **Churn**. This variable records the customers who left the platform within the last month.

The level **"Yes"** represents the customer who has left, while the level **"No"** indicates the customer remains subscribe to our services.


```{r fig.width = 8}
df_churn <- df %>%
  group_by(Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  mutate(Pos = cumsum(Percent)-(Percent/2))

# Churn Distribution (Bar Chart)
p1 <- ggplot(df_churn, aes(Churn, Count), fill = Churn) +
  geom_col(fill = c('#F8766D', '#00BA38')) +
  geom_text(aes(label = Count), vjust = -0.2) +
  ggtitle("Churn Distribution") +
  xlab('Churn') + ylab('Number of Customers') +
  theme_bar


# Churn Rate (Pie Chart)
p2 <- ggplot(df_churn, aes(x = '', y = Percent, fill = Churn)) +
  geom_col(fill = c('#F8766D', '#00BA38')) +
  geom_text(label = sprintf("%.2f%%", df_churn$Percent),
            position = position_stack(vjust = 0.5, reverse = TRUE)) +
  scale_y_continuous(breaks = df_churn$Pos,
                     labels = df_churn$Churn) +
  coord_polar('y') +
  ggtitle("Churn Rate") +
  theme_pie

plot_grid(p1, p2)

```


Among the total number of 7,043 customers in our sample, the ratio of churn to retention is roughly 1:3. This means that out of 10 customers, there are 2~3 of them will likely left.

As we further observe, the churn rate within the last month is pretty high which is around **26%**.

The ideal churn rate is 5% or below. If it reach 10% and above, this signals us to further improve our customer experience in order to retain our customers.


## 4.2. Categorical Variables: Demographic

In order to further understand the distribution of customer churn, we would like to explore more about the demographic data. These customer information includes the gender, age range, marital (partner) status and dependency status.


```{r fig.width = 10}
# Gender plot
df_gender <- df %>%
  group_by(gender, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100)

p3 <- ggplot(df_gender, aes(x = gender, y = Count, fill = Churn)) +
    geom_bar(stat='identity') +
    scale_fill_manual(values=c('#F8766D', '#00BA38')) +
    ggtitle('Gender') +
    theme_bar
    
p4 <- df_gender %>%
  filter(gender=='Female') %>%
  ggplot(aes(x = 2, y = Percent, fill = Churn)) +
    geom_col(fill = c('#F8766D', '#00BA38')) +
    geom_text(aes(x = 2, y = Percent, label = sprintf("%.2f%%", Percent)),
              position = position_stack(vjust = 0.5, reverse = TRUE),
              size = 3) +
    coord_polar('y') + xlim(.2, 2.5) +
    ggtitle("Female") +
    theme_donut

p5 <- df_gender %>%
  filter(gender=='Male') %>%
  ggplot(aes(x = 2, y = Percent, fill = Churn)) +
    geom_col(fill = c('#F8766D', '#00BA38')) +
    geom_text(aes(x = 2, y = Percent, label = sprintf("%.2f%%", Percent)),
              position = position_stack(vjust = 0.5, reverse = TRUE),
              size = 3) +
    coord_polar('y') + xlim(.2, 2.5) +
    ggtitle("Male") +
    theme_donut


p_gender <- plot_grid(p3, plot_grid(p4, p5, ncol = 1),
          get_legend(p3 + theme(legend.position='right')),
          ncol = 3,
          rel_widths = c(1, 1, .25))

# Age Range plot
df_age <- df %>%
  group_by(SeniorCitizen, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100)

p6 <- ggplot(df_age, aes(x = SeniorCitizen, y = Count, fill = Churn)) +
    geom_bar(stat='identity') +
    scale_fill_manual(values=c('#F8766D', '#00BA38')) +
    ggtitle('Senior Citizen') +
    theme_bar

    
p7 <- df_age %>%
  filter(SeniorCitizen=='No') %>%
  ggplot(aes(x = 2, y = Percent, fill = Churn)) +
    geom_col(fill = c('#F8766D', '#00BA38')) +
    geom_text(aes(x = 2, y = Percent, label = sprintf("%.2f%%", Percent)),
              position = position_stack(vjust = 0.5, reverse = TRUE),
              size = 3) +
    coord_polar('y') + xlim(.2, 2.5) +
    ggtitle("No") +
    theme_donut

p8 <- df_age %>%
  filter(SeniorCitizen=='Yes') %>%
  ggplot(aes(x = 2, y = Percent, fill = Churn)) +
    geom_col(fill = c('#F8766D', '#00BA38')) +
    geom_text(aes(x = 2, y = Percent, label = sprintf("%.2f%%", Percent)),
              position = position_stack(vjust = 0.5, reverse = TRUE),
              size = 3) +
    coord_polar('y') + xlim(.2, 2.5) +
    ggtitle("Yes") +
    theme_donut


p_age <- plot_grid(p6, plot_grid(p7, p8, ncol = 1),
          get_legend(p6 + theme(legend.position='right')),
          ncol = 3,
          rel_widths = c(1, 1, .25))

# Partner plot
df_partner <- df %>%
  group_by(Partner, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100)

p9 <- ggplot(df_partner, aes(x = Partner, y = Count, fill = Churn)) +
    geom_bar(stat='identity') +
    scale_fill_manual(values=c('#F8766D', '#00BA38')) +
    ggtitle('Partner') +
    theme_bar

    
p10 <- df_partner %>%
  filter(Partner=='No') %>%
  ggplot(aes(x = 2, y = Percent, fill = Churn)) +
    geom_col(fill = c('#F8766D', '#00BA38')) +
    geom_text(aes(x = 2, y = Percent, label = sprintf("%.2f%%", Percent)),
              position = position_stack(vjust = 0.5, reverse = TRUE),
              size = 3) +
    coord_polar('y') + xlim(.2, 2.5) +
    ggtitle("No") +
    theme_donut

p11 <- df_partner %>%
  filter(Partner=='Yes') %>%
  ggplot(aes(x = 2, y = Percent, fill = Churn)) +
    geom_col(fill = c('#F8766D', '#00BA38')) +
    geom_text(aes(x = 2, y = Percent, label = sprintf("%.2f%%", Percent)),
              position = position_stack(vjust = 0.5, reverse = TRUE),
              size = 3) +
    coord_polar('y') + xlim(.2, 2.5) +
    ggtitle("Yes") +
    theme_donut


p_partner <- plot_grid(p9, plot_grid(p10, p11, ncol = 1),
          get_legend(p9 + theme(legend.position='right')),
          ncol = 3,
          rel_widths = c(1, 1, .25))

# Dependent Plot
df_dependent <- df %>%
  group_by(Dependents, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100)

p12 <- ggplot(df_dependent, aes(x = Dependents, y = Count, fill = Churn)) +
    geom_bar(stat='identity') +
    scale_fill_manual(values=c('#F8766D', '#00BA38')) +
    ggtitle('Dependents') +
    theme_bar

    
p13 <- df_dependent %>%
  filter(Dependents=='No') %>%
  ggplot(aes(x = 2, y = Percent, fill = Churn)) +
    geom_col(fill = c('#F8766D', '#00BA38')) +
    geom_text(aes(x = 2, y = Percent, label = sprintf("%.2f%%", Percent)),
              position = position_stack(vjust = 0.5, reverse = TRUE),
              size = 3) +
    coord_polar('y') + xlim(.2, 2.5) +
    ggtitle("No") +
    theme_donut

p14 <- df_dependent %>%
  filter(Dependents=='Yes') %>%
  ggplot(aes(x = 2, y = Percent, fill = Churn)) +
    geom_col(fill = c('#F8766D', '#00BA38')) +
    geom_text(aes(x = 2, y = Percent, label = sprintf("%.2f%%", Percent)),
              position = position_stack(vjust = 0.5, reverse = TRUE),
              size = 3) +
    coord_polar('y') + xlim(.2, 2.5) +
    ggtitle("Yes") +
    theme_donut


p_dependent <- plot_grid(p12, plot_grid(p13, p14, ncol = 1),
          get_legend(p12 + theme(legend.position='right')),
          ncol = 3,
          rel_widths = c(1, 1, .25))


plot_grid(p_gender, p_age, p_partner, p_dependent, ncol = 2)

```


Based on the plots above, we can conclude that:

- The sample is almost **evenly distributed** across the gender and partner variables.

- Only **minority** of the customers are senior citizens and have dependents.

- For gender, the churn rate is **almost equal** between male and female.

- However, the churn rate is **slightly lower** for those who have partner.

- Customers who are senior citizens or having dependents constitutes a **lower churn rate** as compared to their opposites.












```{r}
glimpse(df)
```

```{r}
summary(df)
```

```{r}
sapply(df, function(x) sum(is.na(x)))
```


```{r}
df$SeniorCitizen <- as.factor(ifelse(df$SeniorCitizen==1, 'Yes', 'No'))
```



```{r fig.width = 10}
# Gender
p1 <- df %>%
  group_by(gender) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x='', y=Percent, fill=gender)) +
    geom_col(fill = c('#2F7FE6', '#F28CFD')) +
    geom_text(aes(x = '', y = cumsum(Percent)-(Percent/2), label = sprintf("%.2f%%", Percent), color=gender),
              show.legend=TRUE, hjust = -0.1, vjust = 1.5, size = 3) +
    coord_polar("y", start = 90) +
    ggtitle("Distribution by Gender") +
    theme_void() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5)) +
    scale_color_manual(values = c("blue","red"))




# SeniorCitizen 
p2 <- df %>%
  group_by(SeniorCitizen) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x='', y=Percent, fill=SeniorCitizen)) +
  geom_col(fill = c('#2F7FE6', '#F28CFD')) +
  coord_polar("y", start = 90) +
  geom_text(aes(x = '', y = cumsum(Percent)-(Percent/2), label = sprintf("%.2f%%", Percent), color=SeniorCitizen),
            show.legend=TRUE, hjust = -0.1, vjust = 1.5, size = 3) +
  ggtitle("Distribution by Age Group") +
  theme_void() +
  theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5)) +
  scale_color_manual(values = c("blue","red"))


# Partner
p3 <- df %>%
  group_by(Partner) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x='', y=Percent, fill=Partner)) +
  geom_col(fill = c('#2F7FE6', '#F28CFD')) +
  coord_polar("y", start = 90) +
  geom_text(aes(x = '', y = cumsum(Percent)-(Percent/2), label = sprintf("%.2f%%", Percent), color=Partner),
            show.legend=TRUE, hjust = -0.1, vjust = 1.5, size = 3) +
  ggtitle("Distribution by Marital Status") +
  theme_void() +
  theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5)) +
  scale_color_manual(values = c("blue","red"))

# Dependents
p4 <- df %>%
  group_by(Dependents) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x='', y=Percent, fill=Dependents)) +
  geom_col(fill = c('#2F7FE6', '#F28CFD')) +
  coord_polar("y", start = 90) +
  geom_text(aes(x = '', y = cumsum(Percent)-(Percent/2), label = sprintf("%.2f%%", Percent), color=Dependents),
            show.legend=TRUE, hjust = -0.1, vjust = 1.5, size = 3) +
  ggtitle("Distribution by Dependency Status") +
  theme_void() +
  theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5)) +
  scale_color_manual(values = c("blue","red"))


grid.arrange(p1, p2, p3, p4, ncol = 2)
```


```{r fig.width = 10}
# Gender
p7 <- df %>%
  group_by(gender, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x=gender, y=Count, fill=Churn)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_text(aes(label=Count, vjust = -0.25), position=position_dodge(1)) +
    scale_fill_manual(values=c('#2F7FE6', '#F28CFD')) +
    xlab('Gender') + ylab('Number of Customers') +
    ggtitle('Customer Churn by Gender') +
    theme_cowplot() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))


# SeniorCitizen 
p8 <- df %>%
  group_by(SeniorCitizen, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x=SeniorCitizen, y=Count, fill=Churn)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_text(aes(label=Count, vjust = -0.25), position=position_dodge(1)) +
    scale_fill_manual(values=c('#2F7FE6', '#F28CFD')) +
    xlab('Senior Citizen') + ylab('Number of Customers') +
    ggtitle('Customer Churn by Age Group') +
    theme_cowplot() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))



# Partner
p9 <- df %>%
  group_by(Partner, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x=Partner, y=Count, fill=Churn)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_text(aes(label=Count, vjust = -0.25), position=position_dodge(1)) +
    scale_fill_manual(values=c('#2F7FE6', '#F28CFD')) +
    xlab('Have Partner') + ylab('Number of Customers') +
    ggtitle('Customer Churn by Marital Status') +
    theme_cowplot() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))


# Dependents
p10 <- df %>%
  group_by(Dependents, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x=Dependents, y=Count, fill=Churn)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_text(aes(label=Count, vjust = -0.25), position=position_dodge(1)) +
    scale_fill_manual(values=c('#2F7FE6', '#F28CFD')) +
    xlab('Have Dependents') + ylab('Number of Customers') +
    ggtitle('Customer Churn by Dependency Status') +
    theme_cowplot() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))


grid.arrange(p7, p8, p9, p10, ncol = 2)

```
