---
title: "WQD7005 Data Mining Project"
author: "Adrian Chen"
date: "23 May 2021"
output: html_document
---

# **1.Accessing and Assaying Prepared Data**

##Customer churn commonly can be describe as customer attrition . 
##The churn refers to the rate at which current subcribing customers who purchase your product or service offering end their relationship with you.
##Thus the churn in customer will result in as a loss in revenue telco business

## The datasets that for this project study will be:
## https://www.kaggle.com/blastchar/telco-customer-churn



# **2.Analysis goal**

### The objectives of this project will be to understand customer behaviors and identifying important variables associated with customer churn.

### The classification decision tree model that will demonstrated by using is ID3 Decision tree . 

### Decision tree is analysis a classification method that uses tree-like models of decisions and their possible outcomes.

## "ID3 stands for Iterative Dichotomiser 3

### The project will explore further on how the ID3 Decision Tree approach by 
### selecting a best attribute that yields maximum Information Gain(IG) or minimum Entropy(H) from the Churn Datasets"



### The SEMMA procedure was created by the SAS Institute. The abbreviation SEMMA represents Sample, Explore,Modify, Model, Assess, and alludes to the way toward leading a data mining project.


### The **Sample** stage for the project comprises of inspecting the information by separating a part of an enormous data collection sufficiently large to contain the hidden important information yet little enough to control rapidly. 



# **3.Analysis data**



```{r}
setwd("C:\\Users\\Adrian\\Desktop\\Awesome\\Mining Churn")
getwd()
```


## Lets read the data of Customers Churn
```{r}
RChurn<-read.csv("TelcoCustomer_Churn.csv")
```
## Lets explore the first 8 rows of the dataframe
```{r}
head(RChurn,8)
```
### Lets explore the last 8 rows of the dataframe
```{r}
tail(RChurn,8)
```

#Exploring the dimension of the data
```{r}
dim(RChurn)
```

```{r}
nrow(RChurn)
```


## Loading the important library
library(dplyr)
```{r}
library(plyr)
library(dplyr)
library(tidyr)
library(corrplot)
library(MASS)
library(e1071)
library(ROCR)
library(gplots)
library(pROC)
library(rpart)
```

Load the fundamental libraries for exploring.

```{r}
library(tidyverse) 
library(ggcorrplot)
library(ggplot2)
library(cowplot)
library(gridExtra)
```




```{r}
sapply(RChurn, function(x) sum(is.na(x)))
```


A few conclusion can be made:

- The data consists of 7,043 observations and 21 variables.

- There are only 3 continuous variables which are **tenure**, **MonthlyCharges** and **TotalCharges**.

- There are only 11 rows with missing data on the variable **TotalCharges**.

- The variable **SeniorCitizen** is not in the categorical form.




##  **4. Show the column metadata**

```{r}
colnames(RChurn)
```


Take a look on the data and identify important information for the preprocessing steps later.

```{r}
glimpse(RChurn)
```

```{r}
summary(RChurn)
```


## 5) **Exploring the data source**
 In the SAS SEMMA of **Explore** , this stage further discovery the data by identifying for hidden trends, important information and anomalies in order to acquire comprehension and insight
 
## **Exploring the data source**
### (creating a histogram, changing the graph properties for a histogram,
### adding a "Missing" Bin to a histogram, creating a pie chart, 
### exploring variable associations)

First, let's see the distribution of our target variable, the **Churn**.


#NOTE: Updated
The variable represents whether the customer will unsubscribe (Yes) or remain subscribe (No) to our services.

```{r}
df<-RChurn
```

```{r fig.width = 8}
df_churn <- df %>%
  # Get the total numbers of churn customers
  group_by(Churn) %>%
  summarise(Count = n()) %>%
  # Calculate the churn rate
  mutate(Percent = prop.table(Count)*100)

# Pie Chart
p1 <- ggplot(df_churn, aes(x='', y=Percent, fill=Churn)) +
  geom_col(fill = c('#2F7FE6', '#F28CFD')) +
  coord_polar("y", start = 90) +
  geom_text(aes(x = 1, y = cumsum(Percent)-(Percent/2), label = sprintf("%.2f%%", Percent)), hjust = -0.1, vjust = 1.5, size = 4.5) +
  ggtitle("Churn Rate") +
  theme_void() +
  theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))

# Bar Chart
p2 <- ggplot(df_churn, aes(Churn, Count), fill=Churn) +
  geom_col(fill = c('#2F7FE6', '#F28CFD'), height = 0.5) +
  geom_text(aes(label = Count), vjust = -0.5, size = 4.5) +
  xlab('Churn') + ylab('Number of Customers') +
  ggtitle("Churn Distribution") +
  theme_cowplot() +
  theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))


grid.arrange(p1, p2, ncol = 2)

```

Based on the churn rate chart, there is about a quarter (26.58%) of our customers churning in the last month.


### 4.3.2. Customer demographic

First, before we would like to know more about demographic data.

The bar chart will be used to compare the population of customers in different gender, age range, partner and dependents.





```{r fig.width = 10}
# Gender
p7 <- df %>%
  group_by(gender, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x=gender, y=Count, fill=Churn)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_text(aes(label=Count, vjust = -0.25), position=position_dodge(1)) +
    scale_fill_manual(values=c('#2F7FE6', '#F28CFD')) +
    xlab('Gender') + ylab('Number of Customers') +
    ggtitle('Customer Churn by Gender') +
    theme_cowplot() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))


# SeniorCitizen 
p8 <- df %>%
  group_by(SeniorCitizen, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x=SeniorCitizen, y=Count, fill=Churn)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_text(aes(label=Count, vjust = -0.25), position=position_dodge(1)) +
    scale_fill_manual(values=c('#2F7FE6', '#F28CFD')) +
    xlab('Senior Citizen') + ylab('Number of Customers') +
    ggtitle('Customer Churn by Age Group') +
    theme_cowplot() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))



# Partner
p9 <- df %>%
  group_by(Partner, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x=Partner, y=Count, fill=Churn)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_text(aes(label=Count, vjust = -0.25), position=position_dodge(1)) +
    scale_fill_manual(values=c('#2F7FE6', '#F28CFD')) +
    xlab('Have Partner') + ylab('Number of Customers') +
    ggtitle('Customer Churn by Marital Status') +
    theme_cowplot() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))


# Dependents
p10 <- df %>%
  group_by(Dependents, Churn) %>%
  summarise(Count = n()) %>%
  mutate(Percent = prop.table(Count)*100) %>%
  ggplot(aes(x=Dependents, y=Count, fill=Churn)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_text(aes(label=Count, vjust = -0.25), position=position_dodge(1)) +
    scale_fill_manual(values=c('#2F7FE6', '#F28CFD')) +
    xlab('Have Dependents') + ylab('Number of Customers') +
    ggtitle('Customer Churn by Dependency Status') +
    theme_cowplot() +
    theme(plot.title = element_text(size = 14, vjust = 2, hjust = 0.5))


grid.arrange(p7, p8, p9, p10, ncol = 2)

```


As observed from the demographic plots above, we can conclude that:

- The distribution and churn rate of the male and female customers is mostly even.

- The distribution of customers based on different partner status is nearly equal.

- The churn rate of those customers who do not have partner is slightly higher.

- Only minority of our customers are senior citizens and having dependents.




## 6) **Modifying and Correcting Source Data**

###In our Data preprocessing, which also part of SAS SEMMMA  **Modify** stage, where involvement of cleaning data,removing redundant attribute which will help for the modelling section and even encoding data for machine learning applications

## To check for missing data
```{r}
colSums(is.na(RChurn))
```

### From the NA data missing data checking we found that the only column that have
### Missing data will be on the TotalCharges with total count on 11 missing 

### There two ways to deal with the missing data, either by mean imputation or removing 
### the missing data completely

### As observed from the **Churn** variable of these affected rows, all these customers are still subscribe to our service.

### Thus,further check on the significance of these rows if removing can be done.

```{r}
sum(is.na(RChurn$TotalCharges))/nrow(RChurn)*100
```
Since this subset constitutes only 0.16% of our data and is quite small, we will just drop them to ease our further analysis task.


```{r}
dim(RChurn)
```

```{r}
RChurn2<-na.omit(RChurn)
dim(RChurn2)
```


## Details of the each variables of the Churn datasets

## Clean and remove Customer ID since no analysis will be done on the column

### As per done earlier the remove irrelevant variables

The variable **customerID** is removed since it will not be used for visualization or modeling.


```{r}
RChurn2$customerID<-NULL
```


## New dimension of of the data will be 7032 rows and 20 Columns variables instead of 21 

```{r}
head(RChurn2)
```

```{r}
dim(RChurn2)
```

```{r}
str(RChurn2)
```

## Using dplyr glimpse gives similar understanding of our variables as well 
glimpse(RChurn)

## summary of the Churn set

```{r}
summary(RChurn2$Churn)
```

## From there we know which will be our interger or double type and which will be our factors

## probailites of Churn
```{r}
prop.table(table(RChurn2$Churn))
```

## Modify columns
##The MultipleLines variable is dependent on the PhoneService variable, where a ‘no’ ##for the latter variable automatically means a ‘no’ for the former variable. We can ##again further ease our graphics and modeling by recoding the ‘No phone service’ ##response to ‘No’ for the MultipleLines variable.

```{r}
```{r}
RChurn2$MultipleLines <-as.factor(plyr::mapvalues(RChurn2$MultipleLines, 
                                          from=c("No phone service"),
                                          to=c("No")))

``` 

RChurn2$MultipleLines 

```{r}
for(i in 9:14){
  RChurn2[,i] <- as.factor(plyr::mapvalues(RChurn2[,i],
                                  from= c("No internet service"), to= c("No")))}

```

##```{r}
##library(dplyr)
##factorrenames <- names(RChurn2[9:14])

## RChurn3 <- RChurn2 %>% mutate_at(.vars=factorrenames, .funs=~recode_factor(., `No ## internet service`="No")) %>% ##mutate_at(.vars="MultipleLines",.funs=~recode_factor(., `No phone service`="No"))

##```



```{r}
RChurn2$MultipleLines

```


### After modification as we observe MultipleLines will be a 2 Factor attributes
```{r}
class(RChurn2$MultipleLines)
str(RChurn2$MultipleLines)


```


### Encoding the attributes of the datasets 

Next, the variable **SeniorCitizen** is encoded into categorical form.

The values are coded in which '0' represents 'No', while the value '1' represents 'Yes'.
```{r}
str(RChurn2)
```

```{r}
glimpse(RChurn2)
```

### From the glimpse of data, the attributes that identify as Factors would be :
###gender ;To encode Female= 0 and Male =1
### Since SeniorCitizen is already integer 1 or 0 will need factor it
### Partner ; Yes =1 No =0
### Dependents : Yes=1 No =0
### PhoneService; Yes=1 No =0
### MultipleLines ; Yes=1 No =0 No phone service=2
### InternetService; DSL=1 , Fiber optic=2 , No=3
### OnlineSecurity ; Yes=1 No =0
### OnlineBackup ;Yes=1 No =0 No phone service=2
### DeviceProtection;Yes=1 No =0 No phone service=2
### TechSupport ; Yes=1 No =0 No phone service=2
### StreamingTV ; Yes=1 No =0 No phone service=2
### StreamingMovies ; Yes=1 No =0 No phone service=2
### Contract : Month-to-month=1 , One year=2 ,Two year=3
### PaperlessBilling ;Yes=1 No =0
### PaymentMethod ; Electronic check=1 Mailed check=2,Bank transfer (automatic)=3
### Credit card (automatic)=4
### Churn=  Yes=1 No =0 




###gender ;To encode Female= 0 and Male =1
```{r}
RChurn2$gender<- factor(RChurn2$gender,
                         levels = c("Female","Male"),
                         labels = c(0,1))
```

### Since Senior Citizen is already integer 1 or 0 will need factor it
```{r}
RChurn2$SeniorCitizen<- as.factor(RChurn2$SeniorCitizen)
```

### Partner ; Yes =1 No =0
```{r}
RChurn2$Partner<- factor(RChurn2$Partner,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```

### Dependents : Yes=1 No =0

```{r}
RChurn2$Dependents<- factor(RChurn2$Dependents,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```

### PhoneService; Yes=1 No =0
```{r}
RChurn2$PhoneService<- factor(RChurn2$PhoneService,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```

### MultipleLines ; Yes=1 No =0 No phone service=2

```{r}
RChurn2$MultipleLines<- factor(RChurn2$MultipleLines,
                        levels = c("No","Yes"),
                        labels = c(0,1))
```


### InternetService; DSL=1 , Fiber optic=2 , No=3
```{r}
RChurn2$InternetService<- factor(RChurn2$InternetService,
                         levels = c("DSL","Fiber optic","No"),
                         labels = c(1,2,3))
```

### OnlineSecurity ; Yes=1 No =0
```{r}
RChurn2$OnlineSecurity<- factor(RChurn2$OnlineSecurity,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```

### OnlineBackup ;Yes=1 No =0 No phone service=2
```{r}
RChurn2$OnlineBackup<- factor(RChurn2$OnlineBackup,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```

### DeviceProtection;Yes=1 No =0 No phone service=2
```{r}
RChurn2$DeviceProtection<- factor(RChurn2$DeviceProtection,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```

### TechSupport ; Yes=1 No =0 No phone service=2
```{r}
RChurn2$TechSupport<- factor(RChurn2$TechSupport,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```
### StreamingTV ; Yes=1 No =0 No phone service=2
```{r}
RChurn2$StreamingTV<- factor(RChurn2$StreamingTV,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```

### StreamingMovies ; Yes=1 No =0 No phone service=2
```{r}
RChurn2$StreamingMovies<- factor(RChurn2$StreamingMovies,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```
### Contract : Month-to-month=1 , One year=2 ,Two year=3
```{r}
RChurn2$Contract<- factor(RChurn2$Contract,levels = c("Month-to-month","One year","Two year"),labels = c(1,2,3))
```

str(RChurn2$Contract)

### PaperlessBilling ;Yes=1 No =0
```{r}
RChurn2$PaperlessBilling<- factor(RChurn2$PaperlessBilling,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```

### PaymentMethod ; Electronic check=1 Mailed check=2,Bank transfer (automatic)=3
### Credit card (automatic)=4

```{r}
RChurn2$PaymentMethod<- factor(RChurn2$PaymentMethod,
                        levels = c("Electronic check","Mailed check",
                        "Bank transfer (automatic)","Credit card (automatic)"),
                        labels = c(1,2,3,4))
```


RChurn2$MultipleLines

### Churn=  Yes=1 No =0 
```{r}
RChurn2$Churn<- factor(RChurn2$Churn,
                         levels = c("No","Yes"),
                         labels = c(0,1))
```



### In this project, data preprocessing technique involving Data Discretization have been performed . The numeric attributes like tenure, MonthlyCharges and TotalCharges will be discretize  for machine learning classification study. The Simple Discretization: Binning of technique like Equal-width (distance) partitioning
have been performed in our study of telco customer Churn datasets.

## Binning can also be used as a discretization technique. Here discretization refers to the process of converting or partitioning continuous attributes, features or variables to discretized or nominal attributes/features/variables/intervals


## For Equal Width Binning,the algorithm function by dividing the the data into k-intervals where each of it have **equal size**. 


##Discretization for tenure
```{r}
RChurn2$tenure <- cut(RChurn2$tenure,7,include.lowest=TRUE, labels=c("0-11", "12-23", "24-35","36-47","48-59","60-71", "72 & above"))
```

```{r}
str(RChurn2$tenure)
```
##Discretization for MonthlyCharges
```{r}
RChurn2$MonthlyCharges <- cut(RChurn2$MonthlyCharges,3,include.lowest=TRUE, labels=c("0-39", "40-79", "80-120"))                   
```


##Discretization for TotalCharges
```{r}
str(RChurn2$TotalCharges)
RChurn2$TotalCharges<- as.numeric(RChurn2$TotalCharges)
RChurn2$TotalCharges <- cut(RChurn2$TotalCharges,8,include.lowest=TRUE, labels=c("0-1086", "1087-2171","2172-3257","3258-4342","4343-5428","5429-6513","6514-7599","7600-8685")) 
```




```{r}
head(RChurn2,3)
```

```{r}
str(RChurn2)
```
